### 2013-06-10 doc/dissertation/document/chapters/parallel_mc/parallel_mc.tex
adding serial verification data for the eigenvalue
restructuring verification results

\chapter{Parallel Monte Carlo Solution Methods for Linear Systems}
\label{ch:parallel_methods}
For Monte Carlo methods, and in particular MCSA, to be viable at the
production scale, scalable parallel implementations are required. In a
general linear algebra context for discrete systems, this has not yet
been achieved. Therefore, we will develop and implement parallel
algorithms for these methods leveraging both the knowledge gained from
the general parallel implementations of Krylov methods and modern
parallel strategies for Monte Carlo as developed by the reactor
physics community. In order to formulate a parallel MCSA algorithm, we
recognize that the algorithm occurs in two stages, an outer iteration
performing fixed point iteration and applying the correction, and an
inner Monte Carlo solver that is generating the correction via the
adjoint or forward methods. The parallel aspects of both these
components must be considered.

In this chapter we briefly review particle transport methods for
domain decomposed Monte Carlo and then provide an analytic framework
from which to estimate its performance. We then devise an algorithm
based on these methods for efficient parallel solutions using the
adjoint Neumann-Ulam method and MCSA and perform parallel scaling
studings to test its performance and compare this performance to
conventional methods.

%%---------------------------------------------------------------------------%%
\section{Domain Decomposition for Monte Carlo}
\label{sec:msod}
As observed in the discussion on parallel Krylov methods, large-scale
problems will surely have their data partitioned such that each
parallel process owns a subset of the equations in the linear
system. Given this convention, the adjoint Monte Carlo algorithm must
perform random walks over a domain that is decomposed and must remain
decomposed due to memory limitations. This naturally leads us to seek
parallel Monte Carlo algorithms that handle domain decomposition.

@@

>>>>

\chapter{Parallel Monte Carlo Synthetic Acceleration}
\label{ch:parallel_methods}
For Monte Carlo methods, and in particular MCSA, to be viable at the
production scale, scalable parallel implementations are required. In a
general linear algebra context for discrete systems, this has not yet
been achieved. Therefore, we will develop and implement parallel
algorithms for these methods leveraging both the knowledge gained from
the general parallel implementations of Krylov methods and modern
parallel strategies for Monte Carlo as developed by the reactor
physics community. In order to formulate a parallel MCSA algorithm, we
recognize that the algorithm occurs in two stages, an outer iteration
performing fixed point iteration and applying the correction, and an
inner Monte Carlo solver that is generating the correction via the
adjoint or forward methods. The parallel aspects of both these
components must be considered.

In this chapter we briefly review particle transport methods for
domain decomposed Monte Carlo and then provide an analytic framework
from which to estimate its performance. We then devise an algorithm
based on these methods for efficient parallel solutions using the
Neumann-Ulam method and MCSA and verify its correctness by comparing
the numerical results against production Krylov methods. We then
perform parallel scaling studings to test its performance on both a
small computing cluster and a leadership class machine and compare
this performance to the production Krylov methods.

%%---------------------------------------------------------------------------%%
\section{Domain Decomposition for Monte Carlo}
\label{sec:msod}
As observed in the discussion on parallel Krylov methods in
Appendix~\ref{ch:linear_problem}, large-scale problems will surely
have their data partitioned such that each parallel process owns a
subset of the equations in the linear system. Given this convention,
the adjoint Monte Carlo algorithm must perform random walks over a
domain that is decomposed and must remain decomposed due to memory
limitations. This naturally leads us to seek parallel Monte Carlo
algorithms that handle domain decomposition.

@@

%%---------------------------------------------------------------------------%%
\section{Parallel MCSA Verification}
\label{sec:parallel_verification}

### 2013-06-10 doc/dissertation/document/chapters/spn_equations/spn_equations.tex
restructuring verification results

\chapter{Monte Carlo Solution Methods for the Simplified $P_N$ Equations}
\label{ch:spn_equations}

@@

For each calculation, the scalar flux in each energy group was
compared over the entire spatial with the GMRES solution used as the
reference.

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/spn_equations/gmres_mcsa_1_1_diff.pdf}
  \end{center}
  \caption{\textbf{Absolute difference of the group 0 scalar flux
      between the MCSA and Aztec GMRES calculations for the 1-group
      calculation.} \textit{Over the entire domain a maximum absolute
      difference of \sn{9}{-10} between the scalar flux computed by
      MCSA and GMRES was observed. The solvers were converged to a
      tolerance of \sn{1}{-8}.}}
  \label{fig:gmres_mcsa_1_1_diff}
\end{figure}

%%---------------------------------------------------------------------------%%
\section{MCSA Performance Comparison to Conventional Methods}
\label{sec:spn_comparison}
Using the same set of solver parameters used for verification, MCSA
will be directly compared to the production Krylov solvers in terms of
both iterative performance and CPU timing. In addition to the
parameters used for the verification analysis, artificial absorption
was also introduced with a value of 0.2 for the MCSA calculations that
used the collision esitmator as CPU timing performance was improved
without a loss in iterative performance. Adding this absorption did
not change the numerical results of the calculation.

>>>>

\chapter{Monte Carlo Synthetic Acceleration Methods for the $SP_N$ Equations}
\label{ch:spn_equations}

@@

For each converged eigenvalue calculation, the scalar flux in each
energy group was compared over the entire spatial with the
BiCGStab-ILUT solution used as the reference to compute an absolute
difference. For the 1 group calculations, these absolute differences
are plotted spatially along the centerline of the fuel assembly in
Figure~\ref{fig:serial_1_group_diff}. Interestingly, except for the
MCSA-ILUT-MR-C calculation, all other 1-group calculations including
GMRES had a maximum global flux difference of \sn{9.313}{-10} even
though different spatial results were observed. For the MCSA-ILUT-MR-C
calculation, a slightly larger maximum error of \sn{2.205}{-9} was
observed for reasons that remain unexplained. All 1-group calculations
have a minimum observed absolute difference of 0 with respect to the
BiCGStab calculation.

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/spn_equations/serial_1_group.pdf}
  \end{center}
  \caption{\textbf{Absolute difference of the group 0 scalar flux
      between the Aztec BiCGStab calculations and Aztec GMRES and the
      MCSA variations for the 1-group calculation.} \textit{Left to
      right: GMRES-ILUT, MCSA-ILUT-R-C, MCSA-ILUT-MR-C,
      MCSA-ILUT-R-EV, MCSA-ILUT-MR-EV.}}
  \label{fig:serial_1_group_diff}
\end{figure}

Table~\ref{tab:serial_differences} gives the maximum and minimum
absolute differences for all calculations and all energy groups using
the BiCGStab calculation as the reference. All comparisons were
performed within a group such that the MCSA-ILUT-R-C results in group
3 for the 4-group calculation were computed with the BiCGStab-ILUT
solution in group 3 for the 4-group calculation as a reference.

%%---------------------------------------------------------------------------%%
\section{MCSA Performance Comparison to Conventional Methods}
\label{sec:spn_comparison}
Using the same set of solver parameters used for verification, MCSA
will be directly compared to the production Krylov solvers in terms of
both iterative performance and CPU timing for solutions to the fuel
assembly problem. In addition to the parameters used for the
verification analysis, artificial absorption was also introduced with
a value of 0.2 for the MCSA calculations that used the collision
esitmator as CPU timing performance was improved without a loss in
iterative performance. Adding this absorption did not change the
numerical results of the calculation.
